---
title: "Data Analysis Biomas Traits"
author: "Md Neaz Ali"
date: "2024-07-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data and Library Import
```{r message=FALSE, error=FALSE}
library(agricolae)
library(dendextend)
library(dplyr)
library(ggplot2)
library(pheatmap)
library(metan)
library(stats)
library(MASS)
library(factoextra)
library(gridExtra)
library(gt)
library(psych)
library(ggdendro)
library(tidyverse)
library(lme4)
library(lsmeans)
```
## Importing Data
```{r}
mva_data <- read.csv("./data/mva.csv",header = TRUE)
anova_data <- read.csv("./data/anova.csv", header = TRUE)
# initial_data <- read.csv("./data/initial.csv, header = TRUE")
boxplot_data <- read.csv("./data/boxplot.csv", header = TRUE)
```

# ANOVA
```{r}
anova_data %>% 
  str()
anova_data$Rep <- as.factor(anova_data$Rep)
anova_data$Treat <- as.factor(anova_data$Treat)
anova_data$Geno <- as.factor(anova_data$Geno)
anova_data %>% 
  str()
```
## Summary
```{r}
summary(anova_data[,-c(1,2,3)])
```
## Standard Deviations
```{r}
variables <- names(anova_data)
for(i in 4:11){
  output <- paste0(variables[i]," : ", sd(anova_data[,i]) %>% round(3))
  print(output)
}
```
## anova
```{r}
variables <- names(anova_data)
variables <- variables[4:11]
for(var in variables){
  print(paste0("Anova for: ", var))
  formula <- as.formula(paste(var, "~Rep+Treat+Geno"))
  model <- aov(formula, data = anova_data)
  print("anova result : ------------------------")
  print(summary(model))
}
```


# Boxplots
```{r}
box_data <- read.csv("./data/Boxplot_edited2.csv", header = T)
str(box_data)
```
```{r}
box_data$Geno <- as.factor(box_data$Geno)
box_data$Rep <- as.factor(box_data$Rep)
box_data$Treat <- as.factor(box_data$Treat)
```


```{r}
variables <- names(box_data)[4:11]
path <- getwd()
output_dir <- paste0(path, "/images/boxplots/")
if(!dir.exists(output_dir)){
  dir.create(output_dir, recursive = TRUE)
}
```

for(var in variables){
  formula <- as.formula(paste0(var ,"~ Treat"))
  print(paste("Boxplot for : ", var))
  file_name_box <- paste0(output_dir,var ,".png")
  jpeg(filename = file_name_box, width = 900, height = 600, quality = 600)
  par(mar = c(5,4,4,2)+0.3)
  boxplot(formula, col = 5, data = box_data)
  dev.off()
  boxplot(formula, col = 5, data = box_data)
}


```{r}
for (var in variables) {
  formula <- as.formula(paste0(var, " ~ Treat"))
  print(paste("Boxplot for: ", var))
  
  p <- ggplot(box_data, aes_string(x = "Treat", y = var)) +
    geom_boxplot(fill = "#8cbf13", color = "black", outlier.color = "red", outlier.shape = 16, outlier.size = 2) +
    labs(x = " ", y = var) +
    theme_minimal(base_size = 15) +  # Use a minimal theme with larger base font size
    theme(
      axis.text.x = element_text(hjust = 1), 
      panel.background = element_rect(fill = "white"),
      plot.background = element_rect(fill = "white"),  
      panel.grid.major = element_blank(),  
      panel.grid.minor = element_blank()  
    )
  
 
  file_name_box <- paste0(output_dir, var, ".png")
  ggsave(file_name_box, plot = p, width = 9, height = 6, dpi = 300)
  p
}

```

# PCA 
```{r}
mva_data %>% 
  head()
```
```{r}
mva_data %>% 
  str()
```
```{r}
mva_data$Rep <- as.factor(mva_data$Rep)
mva_data$Geno <- as.factor(mva_data$Geno)
```

### pca calculations
```{r}
pca <- prcomp(mva_data[,-c(1,2)], center = TRUE, scale = TRUE)
print(summary(pca))
```
```{r}
print(names(pca))
```

```{r}
pca$rotation
```

```{r}
pca$x
```
```{r}
principal_comps <- pca$x
correlations <- cor(mva_data[,-c(1,2)], principal_comps)
print(correlations)
```


### Scree Plot of Variance
```{r}
eigenvalues <- pca$sdev^2
explained_variance <- eigenvalues/sum(eigenvalues)*100

scree_data <- data.frame(
  PC = paste0("PC", 1:length(explained_variance)),
  variance = explained_variance
)

hline_value = 3
scree_plot <- ggplot(scree_data, aes(x = PC, y = variance))+
  geom_bar(stat = "identity", fill = "#8cbf13")+
  geom_text(aes(label = round(variance,1)), vjust = -0.5)+
  geom_line(aes(x = 1:length(pca$sdev),y = pca$sdev^2/sum(pca$sdev^2)*100),color = "blue", size =0.4 )+
  geom_hline(yintercept = hline_value,linetype = "dashed", color = "red",lwd = 0.4)+
  xlab("Principal Component")+
  ylab("Percentage of variance Explained")+
  ggtitle("Scree Plot")
ggsave("./images/scree_plot/scree_plot.png", scree_plot,width = 12, height = 8, dpi = 600)
scree_plot
```
### Biplot
```{r}
biplot <- fviz_pca_biplot(pca,
                          geom.ind = "point",
                          pointshape = 21,
                          pointsize = 2,
                          fill.ind = mva_data$Geno,
                          col.ind = "black",
                          col.var = "blue",
                          legend.title = "Genotypes",
                          repel = TRUE)
biplot <- biplot +
  geom_point(aes(x = pca$x[,1], y = pca$x[,2], shape = mva_data$Geno, color= mva_data$Geno), size=2, stroke = 1.5)
biplot
```

# Clustering
### Dendogram

```{r}
# Step 1: Prepare the Data
clust_data <- mva_data[, -1]

# Step 2: Aggregate Data by Genotype
clust_data_agg <- clust_data %>%
  group_by(Geno) %>%
  summarise(across(where(is.numeric), mean, na.rm = TRUE))

# Step 3: Standardize the Numerical Data  
clust_data_scaled <- clust_data_agg %>%
  mutate(across(where(is.numeric), scale))

# Step 4: Calculate the Distance Matrix
dist_matrix <- dist(clust_data_scaled, method = "euclidean")

# Step 5: Perform Hierarchical Clustering
hc <- hclust(dist_matrix)

# Step 6: Convert to dendrogram data
dendro_data <- ggdendro::dendro_data(hc)

# Create the dendrogram plot
dendrogram_plot <- ggplot(width = 12) +  # Adjust the width of the plot
  geom_segment(data = dendro_data$segments, aes(x = x, y = y, xend = xend, yend = yend), size = .4, color = "#8cbf13") +
  geom_text(data = dendro_data$labels, aes(x = x, y = y, label = label), hjust = -0.1, size = 3) +
  geom_hline(yintercept = 8, linetype = "dashed", color = "blue", size = .3)+
  theme_minimal() +
  labs(title = "Dendrogram Using Agglomerative Clustering Method", x = "", y = "Height") +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())
ggsave("./images/dendogram/dendogram_plot.png", plot = dendrogram_plot, width = 12, height = 8, dpi = 600)
dendrogram_plot
```


### Cluster Members
```{r}
clust_data <- mva_data[, -1]
clust_data_agg <- clust_data %>% 
  group_by(Geno) %>% 
  summarise(across(where(is.numeric), mean, na.rm = TRUE))
clust_data_scaled <- clust_data_agg %>%
  mutate(across(where(is.numeric), scale))

dist_matrix <- dist(clust_data_scaled, method = "euclidean")
hc <- hclust(dist_matrix)
num_clusters <- 4
cluster_membership <- cutree(hc, k = num_clusters)
clust_data_agg <- clust_data_agg %>%
  mutate(Cluster = cluster_membership)

table(clust_data_agg$Geno, clust_data_agg$Cluster)
```

```{r}
clust_data_agg$Cluster <- as.factor(clust_data_agg$Cluster)
head(clust_data_agg)
```

# Mean values of four clusters based on quantitative traits
```{r}
groups <- c(1,2,3,4)
for(i in groups){
  summ <- clust_data_agg[,-1] %>% 
            filter(Cluster == i) %>% 
            summary()
  print(paste0("Summary of Cluster number: ",i))
  print(summ)
}
```


# Line Tester Mating 
```{r}
head(anova_data)
```
```{r}
model <- lmer(LN~Geno+(1|Rep)+(1|Geno:Rep),data = anova_data)
```
```{r}
summary(model)
```
```{r}
VarCorr(model)
```
Prdicting the values
```{r}
predicted <- predict(model)
anova_data$predicted <- predicted
```

GCA and SCA Calculation
```{r}
gca <- anova_data %>% 
  group_by(Geno) %>% 
  summarise(GCA = mean(predicted))
sca <- anova_data %>% 
  group_by()
```


```{r}
lsm <- lsmeans(model, ~Geno)
summary(lsm)
```
```{r}
sink("linextesteroutput.txt")
# with(anova_data,lineXtester(Rep, Geno, LN))
sink()
```

# GV, AV, DV etc
```{r}
gvdat <- read.csv("./data/GV_data.csv")
gvdat %>% head()
```
```{r}
str(gvdat)
```


```{r}
df <- gvdat$df
Source <- gvdat$Source
rownames(gvdat) <- Source
gvdat <- gvdat[,-c(1,2)]
gvdat %>%  head()
```
```{r}
head(anova_data)
```
```{r}
means <- data.frame(Criterias = c("LN", "Rl",	"SL",	"LA",	"SDW",	"RDW",	"LDW",	"TDM" ), 
           means = c(6.38,	14.82,	8.318,	52.4,	0.05568,	0.04149,	0.1254,	0.2226))
```


```{r}
GV <- (gvdat["Treatment",] - gvdat["Error",])/4
GV
```

```{r}
EV <- gvdat["Error",]
EV
```

```{r}
PV <- GV + EV
PV
```
```{r}
GCV <- (GV/means$means) * 100
GCV
```
```{r}
PCV <- (PV/means$means) * 100
PCV
```
```{r}
H.broad_sense <- GV/ PV
H.broad_sense
```
```{r}
H.percent <-  H.broad_sense * 100
H.percent
```
```{r}
GA <- (H.broad_sense^2) * 2.06 * sqrt(PV)
GA
```






